---
title: "Prelim analyses"
author: "CLM"
date: "5/9/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a walk through of how to analyze the complete dataset for insect presence in Baptisia fruits from along the Allegheny River. 

It's part tutorial, part test code. 


```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("MASS")) install.packages("MASS")
library(MASS)
install.packages("cowplot", repos='http://cran.us.r-project.org')
library(cowplot)
install.packages("UsingR", repos='http://cran.us.r-project.org')
library(UsingR)
library(dplyr)
```


Our data is stored as a cvs, kind of like an excels sheet that R will recognize. Make sure the file in is the same folder as this .Rmd
```{r read in data}
prelim_data<-read.csv("prelim_Bap_fruit_data.csv", sep=",", header=TRUE)
```
It should now be  the Environment pane in the upper righthand side

##Data Exploration
An important part of working in R is making sure that the data you read in looks right
```{r number of columns and rows}
nrow(prelim_data)
ncol(prelim_data)
```
```{r structure of data}
str(prelim_data)
#we may have to come back and change some of the data types, esp b/c I used 0 and 1 for presence absence, so in our case this isn't numeric data it's factor/categorical
```
Data types: 
```{r top and bottom}
head(prelim_data)
tail(prelim_data)
```

We also what to check Ns, whcih means how much data we have and if there are any NAs or empty categories. With this prelim data there are a few.

```{r check Ns}
apply(prelim_data, 2, function(x) 
  round(100 * (length(which(is.na(x))))/length(x) , digits = 1)) %>%
  as.data.frame() %>%
  `names<-`('Percent of Missing Values')
```
To fix the NA problem, there are several approaches remove any observation with an NA of fill it with averages of surrounding values. The best approach here probably to remove the NAs since there are only a few (the ones for immature seeds are the ones we want to remove b/c each ind should have a measure). Hopefully with the full data, we won't have any NAs to worry about. 


##Tidy data and data manipulation
You may have heard of "tidy" data in R. This is just a format for data to make it easy to work with. I tried to design and edit the spreadsheet before making it a csv to make it "tidy." This will allow us to easily manipulate columns etc.
```{r making an avg length column for seeds}
#WHY YOU NO WORK
#prelim_data<- prelim_data %>% mutate(avg_seed_length =rowMeans(10:14), na.rm=T)
#mydata <- cbind(data,IVMean=rowMeans(data[3:8], na.rm=TRUE))
prelim_data<-cbind(prelim_data, avg_seed_length=rowMeans(prelim_data[10:14], na.rm=TRUE))
#prelim_data<- mutate(df, avg_seed_length = rowMeans(select(prelim_data), c(10:14), na.rm = TRUE))
```
https://stackoverflow.com/questions/28744419/dplyr-mean-for-multiple-columns

##Dealing with NAs
```{r replacing NA with fill function}
vars_to_fill <- c(8:14) #this is everything from Plant_damage to Immature_seeds  
no_NA_prelim_data <- prelim_data %>% tidyr::fill (vars_to_fill)
```



#Normal distribution?

```{r distribution of DBH}
qqnorm(prelim_data$Tree.dbh...cm. , main='Normal Q-Q Plot for DBH')
```

Hmmm maybe normal, let's look more closely with the function 'simple.eda()' from the UsingR package.

```{r EDA}
simple.eda(prelim_data$Length_.mm.)
simple.eda(prelim_data$Number_of_racemes)
simple.eda(prelim_data$Fruits_on_plant)
simple.eda(prelim_data$Height_.m.)
simple.eda(prelim_data$Approx_dist_to_Allegheny_.m.)
simple.eda(prelim_data$Number_of_mature_seeds)
simple.eda(prelim_data$Immature_seeds)
simple.eda(prelim_data$Avg_seed_length_.mm.)
```
Probably, maybe? 

```{r shapiro-wilk test }
shapiro.test(prelim_data$Length_.mm.)
shapiro.test(prelim_data$Number_of_racemes)
shapiro.test(prelim_data$Fruits_on_plant)
shapiro.test(prelim_data$Height_.m.)
shapiro.test(prelim_data$Approx_dist_to_Allegheny_.m.)
shapiro.test(prelim_data$Number_of_mature_seeds)
shapiro.test(prelim_data$Immature_seeds)
shapiro.test(prelim_data$Avg_seed_length_.mm.)
```
Not:
Mature seeds
Height
Everything else is normally dist

```{r}
shapiro.test(log10(prelim_data$Number_of_mature_seeds))
#log not great for this, I think the issue is an outlier 
shapiro.test(log(prelim_data$Height_.m.))
#cuts p in half 
```


##Using a generalized linear model (glm)

```{r}
glm<- glm(family=binomial, Damage ~ Site + Length_.mm. +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= no_NA_prelim_data, na.action=na.pass)
summary(glm)
```
I think I need Poisson b/c I have mixed response variables when I use seed as response but damage is y/n so I think binomial
https://personal.psu.edu/abs12/stat504/online/07_poisson/02_poisson_beyond.htm
https://newonlinecourses.science.psu.edu/stat504/node/216/
```{r}
mod.selection<-stepAIC(glm, trace=TRUE, na.action=na.pass) 

summary(mod.selection) 
```

glm(formula = Damage ~ Insect_presence + Frass_presence + Fruit_evidence_of_insect + 
    Number_of_racemes + Number_of_mature_seeds, family = binomial, 
    data = no_NA_prelim_data, na.action = na.pass)
AIC: 12

##Separating impacted and not
```{r}
impacted<- prelim_data %>% filter(Fruit_evidence_of_insect==1) 
unimpacted <- prelim_data %>% filter(Fruit_evidence_of_insect==0) 
```

```{r}
summary(impacted)
summary(unimpacted)
```
avg number mature 
impacted:12.91
unimpacted:13.68

avg immature
impacted: 4
unimpacted:4.128

avg seed length
impacted: 4.961
unimpacted: 5.155

```{r}
fruit_length<- mean(impacted$Length_.mm., na.rm =TRUE)
number_of_fruits<- mean(impacted$Fruits_on_plant, na.rm=TRUE)
mature_seeds<- mean(impacted$Number_of_mature_seeds, na.rm=TRUE)
immature_seeds<- mean(impacted$Immature_seeds, na.rm=TRUE)
seed_length<- mean(impacted$Avg_seed_length_.mm., na.rm=TRUE)
```

```{r}
impacted_avg <- cbind(fruit_length,number_of_fruits, mature_seeds, immature_seeds, seed_length)
```

```{r}
fruit_length2<- mean(unimpacted$Length_.mm., na.rm =TRUE)
number_of_fruits2<- mean(unimpacted$Fruits_on_plant, na.rm=TRUE)
mature_seeds2<- mean(unimpacted$Number_of_mature_seeds, na.rm=TRUE)
immature_seeds2<- mean(unimpacted$Immature_seeds, na.rm=TRUE)
seed_length2<- mean(unimpacted$Avg_seed_length_.mm., na.rm=TRUE)
```

```{r}
unimpacted_avg<- cbind(fruit_length2, number_of_fruits2, mature_seeds2, immature_seeds2, seed_length2)
```

```{r}
dataframe<- rbind(impacted_avg, unimpacted_avg)
```

I don't know how helpful the above is. I took some time to reformat data in excel. 
```{r}
ttest.data<- read.csv("ttest_data.csv", sep=",", header=TRUE) 
```
```{r}
vars_to_fill <- c(2:6)
no.NA.ttest.data<- prelim_data %>% tidyr::fill (vars_to_fill)
```


ttest
```{r}
t.test(Number_of_mature_seeds ~ status, ttest.data)
t.test(Immature_seeds ~ status, ttest.data)
t.test(Avg_seed_length_.mm. ~ status, ttest.data)
t.test(Fruits_on_plant ~ status, ttest.data)
```
https://uc-r.github.io/t_test


###glm for one site
```{r}
site1<- no_NA_prelim_data %>% filter (Site==1)
```

```{r}
glm_site1<- glm(family= binomial, Damage ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= site1, na.action=na.pass)
summary(glm_site1)
```

```{r}
mod.selection_s1<-stepAIC(glm_site1, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s1) 
```


##Site 9

```{r}
site9<- no_NA_prelim_data %>% filter (Site==9)
```

```{r}
glm_site9<- glm(family=binomial, Damage ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= site9, na.action=na.pass)
summary(glm_site9)
```

```{r}
mod.selection_s9<-stepAIC(glm_site9, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s9)
```

##Site 14

```{r}
site14<- no_NA_prelim_data %>% filter (Site==14)
```

```{r}
glm_site14<- glm(family=binomial, Damage ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= site14, na.action=na.pass)
summary(glm_site14)
```

```{r}
mod.selection_s14<-stepAIC(glm_site14, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s14)
```


##Site 21

```{r}
site21<- no_NA_prelim_data %>% filter (Site==21)
```

```{r}
glm_site21<- glm(family=binomial, Damage ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= site21, na.action=na.pass)
summary(glm_site21)
```

```{r}
mod.selection_s21<-stepAIC(glm_site21, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s21)
```


##Site26

```{r}
site26<- no_NA_prelim_data %>% filter (Site==26)
```

```{r}
glm_site26<- glm(family=binomial, Damage ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Number_of_mature_seeds + Immature_seeds + Avg_seed_length_.mm. , data= site26, na.action=na.pass)
summary(glm_site26)
```

```{r}
mod.selection_s26<-stepAIC(glm_site26, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s26)
```


##Graphs to direct next analyses
```{r}
library(ggplot2)
```

```{r}
pop_fruit_data$site <- as.character(pop_fruit_data$site)
```

```{r}
#need to change data type for the site 
damage_graph<-ggplot(pop_fruit_data, aes(x=site, y=no.plant_damage)) +
  geom_bar(stat="identity")
damage_graph
```


##another glm

###seeds
```{r}
glm_seed<- glm(Number_of_mature_seeds ~ Damage+ Site + Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Immature_seeds + Avg_seed_length_.mm. , family=poisson, data= no_NA_prelim_data, na.action=na.pass)
summary(glm_seed)
```

```{r}
mod.selection_seed<-stepAIC(glm_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_seed) 
```

#####individual sites

##Site 1

```{r}
glm_site1_seed<- glm(Number_of_mature_seeds ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Damage + Immature_seeds + Avg_seed_length_.mm. , data= site1, na.action=na.pass)
summary(glm_site1_seed)
```

```{r}
mod.selection_s1_seed<-stepAIC(glm_site1_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s1_seed) 
```
nada

##Site 9

```{r}
glm_site9_seed<- glm(Number_of_mature_seeds ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Damage + Immature_seeds + Avg_seed_length_.mm. , data= site9, na.action=na.pass)
summary(glm_site9_seed)
```

```{r}
mod.selection_s9_seed<-stepAIC(glm_site9_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s9_seed)
```
*fruits on plants
*approx dist to Allegheny 

##Site 14

```{r}
glm_site14_seed<- glm( Number_of_mature_seeds ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Damage + Immature_seeds + Avg_seed_length_.mm. , data= site14, na.action=na.pass)
summary(glm_site14_seed)
```

```{r}
mod.selection_s14_seed<-stepAIC(glm_site14_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s14_seed)
```
*length
*frass
*approx dist to Allegheny

##Site 21

```{r}
glm_site21_seed<- glm( Number_of_mature_seeds ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Damage + Immature_seeds + Avg_seed_length_.mm. , data= site21, na.action=na.pass)
summary(glm_site21_seed)
```

```{r}
mod.selection_s21_seed<-stepAIC(glm_site21_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s21_seed)
```
**length
**insect presence
**immature seeds

##Site26

```{r}
glm_site26_seed<- glm( Number_of_mature_seeds ~ Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Fruits_on_plant + Height_.m. + Approx_dist_to_Allegheny_.m. + Damage + Immature_seeds + Avg_seed_length_.mm. , data= site26, na.action=na.pass)
summary(glm_site26_seed)
```

```{r}
mod.selection_s26_seed<-stepAIC(glm_site26_seed, trace=TRUE, na.action=na.pass) 

summary(mod.selection_s26_seed)
```
nada

###fruits
```{r}
glm_fruits<- glm(Fruits_on_plant~ Damage+ Number_of_mature_seeds+ Site + Length_.mm. + Insect_presence + Frass_presence + Fruit_evidence_of_insect +Plant_damage + Number_of_racemes + Height_.m. + Approx_dist_to_Allegheny_.m. + Immature_seeds + Avg_seed_length_.mm. , data= no_NA_prelim_data, na.action=na.pass)
summary(glm_fruits)
```

```{r}
mod.selection_fruits<-stepAIC(glm_fruits, trace=TRUE, na.action=na.pass) 

summary(mod.selection_fruits) 
```