---
title: "seed count ttest"
author: "CLM"
date: "12/6/2019"
output: html_document
---

This Rmd contains code to do a ttest between number of seeds found along the Allegheny River in PA for my reseach vs the number of seeds reported for the species in keys. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r whole dataframe}
prelim_data<-read.csv("prelim_Bap_fruit_data.csv", sep=",", header=TRUE)
```

```{r seed dataframe}
seed_data<- read.csv("Baptisia_seed_counts.csv", sep=",", header=TRUE)
```

```{r}
str(seed_data)
```
```{r}
seed_data$Site<- as.factor(seed_data$Site)
seed_data$Expected_number_of_seeds<- as.numeric(seed_data$Expected_number_of_seeds)
seed_data$Number_of_seeds<- as.numeric(seed_data$Number_of_seeds)
```

```{r}
seed_data<-na.omit(seed_data)
```

ttest
```{r}
t.test(data= seed_data, Number_of_seeds, Expected_number_of_seeds)
```
So this is isn't working: following error "Error in t.test(data = seed_data, Number_of_seeds, Expected_number_of_seeds) : 
  object 'Number_of_seeds' not found"
  
  
I think I need to do ttest with just averages and do a one way ttest


```{r}
t.test(seed_data$Number_of_seeds, mu = 25, alternative = "two.sided")
```


there ya go! significantly less seeds in Allegheny pops than expected according to Isely

should also check per pop 
```{r}
seeds_site1 <- seed_data %>% filter(Site==1) 
seeds_site9 <- seed_data %>% filter(Site==9) 
seeds_site14 <- seed_data %>% filter(Site==14) 
seeds_site21 <- seed_data %>% filter(Site==21) 
seeds_site26 <- seed_data %>% filter(Site==26) 
```


```{r}
t.test(seeds_site1$Number_of_seeds, mu = 25, alternative = "two.sided")
t.test(seeds_site9$Number_of_seeds, mu = 25, alternative = "two.sided")
t.test(seeds_site14$Number_of_seeds, mu = 25, alternative = "two.sided")
t.test(seeds_site21$Number_of_seeds, mu = 25, alternative = "two.sided")
t.test(seeds_site26$Number_of_seeds, mu = 25, alternative = "two.sided")
```

Yep! Golden

Next I should do ttest for plants w/ no damage

```{r}
prelim_data<-read.csv("prelim_Bap_fruit_data.csv", sep=",", header=TRUE)
```
```{r}
prelim_ttest<- prelim_data %>% filter(Damage==0)
```

```{r}
t.test(prelim_ttest$Number_of_mature_seeds, mu = 25, alternative = "two.sided")
```
still significant




http://www.sthda.com/english/wiki/one-sample-t-test-in-r
https://uc-r.github.io/t_test
